---
title: "General Cell Chat analysis"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: no
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: true
    theme: lumen
params:
  reduction.name: NA
  cluster.name: NA
  path.to.s.obj: NA
  outputdir: NA
---

```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=14, fig.height=10}
library(patchwork)
options(stringsAsFactors = FALSE)
library("CellChat")
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=14, fig.height=10}
scrna_pipeline_src <- "/media/hieunguyen/HNSD01/src/src_pipeline/scRNA_GEX_pipeline_SeuratV5/processes_src"
source(file.path(scrna_pipeline_src, "import_libraries.R"))
source(file.path(scrna_pipeline_src, "helper_functions.R"))

outdir <- "/media/hieunguyen/GSHD_HN01/outdir"
main.PROJECT <- "BrainMet_SeuratV5"

reduction.name <- params$reduction.name
cluster.name <- params$cluster.name
path.to.s.obj <- params$path.to.s.obj
path.to.save.output <- params$outputdir

sample.id <- "all"
filter10cells <- "Filter10"

dir.create(path.to.save.output, showWarnings = FALSE, recursive = TRUE)

s.obj <- readRDS(path.to.s.obj)

```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=14, fig.height=10}
DefaultAssay(s.obj) <- "SCT"
```

# Dataset overview

## All samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, reduction = reduction.name, label = TRUE, label.box = TRUE, repel = TRUE, group.by = cluster.name)
```


# Basic Data analysis
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
##### Get the input data matrix
data.input <- GetAssayData(object = s.obj, slot = "data", assay = "SCT")
# meta.data <- s.obj@meta.data %>% 
#   rownames_to_column("barcode") %>%
#   rowwise() %>%
#   mutate(cluster = sprintf("cluster_%s", cca.cluster.0.5)) %>%
  
meta.data <- s.obj@meta.data %>%
  rownames_to_column("barcode")
meta.data$cluster <- unlist(lapply(meta.data[[cluster.name]], function(x){
  return(sprintf("cluster_%s", x))
}))
meta.data <- meta.data %>% column_to_rownames("barcode")

num.cluster <- length(unique(s.obj@meta.data[[cluster.name]]))

check.cluster.typeof <- sample(unique(s.obj@meta.data[[cluster.name]]), 1)
# if (cluster.name != "cell.annotation"){
if (typeof(check.cluster.typeof) %in% c("double", "integer")){
  all.cluster.idx <- to_vec (for (cluster.id in unique(s.obj@meta.data[[cluster.name]])) str_replace(cluster.id, "cluster_", "")) %>% as.numeric()
  meta.data$cluster <- factor(meta.data$cluster, levels = to_vec ( for (i in sort(all.cluster.idx)) sprintf("cluster_%s", i) ))  
} else {
  meta.data$cluster <- meta.data[[cluster.name]]
}

cellchat <- createCellChat(object = data.input, meta = meta.data, group.by = "cluster")

##### Some preprocessing steps for the cellchat object
cellchat <- addMeta(cellchat, meta = meta.data)
cellchat <- setIdent(cellchat, ident.use = "cluster") # set "labels" as default cell identity
groupSize <- table(cellchat@idents) %>% as.data.frame() # number of cells in each cell group
groupSize %>% create_dt()
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
##### Define the CELLCHAT DATABASE IN USE
CellChatDB <- CellChatDB.mouse
showDatabaseCategory(CellChatDB)
```

This dataframe shows all available interaction in the current CellChat (mouse) database
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
CellChatDB$interaction %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
##### Subset or use everything in the CellChat database
##### In this analysis, we use all available interactions
if (file.exists(file.path(path.to.save.output, sprintf("CellChat_object.%s.%s.rds", sample.id, filter10cells))) == FALSE){
  # CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling") # use Secreted Signaling
  CellChatDB.use <- CellChatDB.human # simply use the default CellChatDB
  
  # Remove the interaction H2-BI and H2-Ea-ps
  # CellChatDB.use[["interaction"]] <- CellChatDB.use[["interaction"]][-which(CellChatDB.use[["interaction"]]$ligand == "H2-BI"),]
  # CellChatDB.use[["interaction"]] <- CellChatDB.use[["interaction"]][-which(CellChatDB.use[["interaction"]]$ligand == "H2-Ea-ps"),]
  
  cellchat@DB <- CellChatDB.use
  
  # subset the expression data of signaling genes for saving computation cost
  cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
  
  ##### Run parallelly
  future::plan("multicore", workers = 10) # do parallel
  
  ##### Preprocessing the expression data matrix for cell-cell communication analysis
  cellchat <- identifyOverExpressedGenes(cellchat)
  cellchat <- identifyOverExpressedInteractions(cellchat)
  
  ##### Compute the communication probability and infer cellular communication network
  cellchat <- computeCommunProb(cellchat)
  if (filter10cells == "Filter10"){
    cellchat <- filterCommunication(cellchat, min.cells = 10)
  }
    
  ##### Infer the cell-cell communication at a signaling pathway level
  cellchat <- computeCommunProbPathway(cellchat)
  
  ##### Calculate the aggregated cell-cell communication network
  cellchat <- aggregateNet(cellchat)
  
  saveRDS(object = cellchat, file = file.path(path.to.save.output, sprintf("CellChat_object.%s.%s.rds", sample.id, filter10cells)))
} else {
  cellchat <- readRDS(file.path(path.to.save.output, sprintf("CellChat_object.%s.%s.rds", sample.id, filter10cells)))
}
```

## Table: Signal between each pair of clusters
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
mat.weight <- cellchat@net$weight
mat.count <- cellchat@net$count

matP.weight <- cellchat@netP$prob
```

### Counts
The number of interactions between any two cell groups  (row --> column, row: source, column: target)
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
mat.count %>% create_dt()
```

### Weights
The total reaction strength - weights. Weights are proportional to the interaction strength.
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
mat.weight %>% create_dt()
```

# Visualization: interactions
## Aggregated interaction network between all clusters {.tabset}

(not really informative!)

### The number of interactions (counts)
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(mat.count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
```

### The strength of interactions (weights)
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
netVisual_circle(mat.weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

## Sent signals from each cell group
Since it is difficult to interpret the interactions on the above aggregated plots, here we show the strength of signals sent from each cluster to the others. Thicker edge line indicates a stronger signal. In the Circle plot below, circle sizes are proportional to the number of cells in each cluster. 

The weights can be compared globally between clusters; for example, if the line connecting cluster 0 and cluster 10 in the plot `cluster 0` is **thicker** than the line connecting cluster 1 and cluster 10, we can say that the sending signal from 0 --> 10 is **stronger** than 1 --> 10.  

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
par(mfrow = c(2, 2), xpd=TRUE)
for (i in 1:nrow(mat.weight)) {
  mat2 <- matrix(0, nrow = nrow(mat.weight), ncol = ncol(mat.weight), dimnames = dimnames(mat.weight))
  mat2[i, ] <- mat.weight[i, ]
  mat2[i, i] <- 0
  strength <- mat2[i, ] %>% sort() %>% tail(5)
  strength <- strength[strength != 0]
  names(strength) <- to_vec(for (item in names(strength)) str_replace(item, "cluster_", ""))
  title <- sprintf("Top 5 highest signals from cluster %s: %s", rownames(mat.weight)[i], paste(names(strength), collapse = ", "))
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat.weight), title.name = title, arrow.size = 1)
}
```

## Received signals from each cell group

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
par(mfrow = c(2, 2), xpd=TRUE)
for (i in 1:nrow(mat.weight)) {
  mat2 <- matrix(0, nrow = nrow(mat.weight), ncol = ncol(mat.weight), dimnames = dimnames(mat.weight))
  mat2[, i] <- mat.weight[, i]
  mat2[i, i] <- 0
  strength <- mat2[, i] %>% sort() %>% tail(5) 
  strength <- strength[strength != 0]
  names(strength) <- to_vec(for (item in names(strength)) str_replace(item, "cluster_", ""))
  title <- sprintf("Top 5 highest signals to cluster %s: %s", rownames(mat.weight)[i], paste(names(strength), collapse = ", "))
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat.weight), title.name = title, arrow.size = 1)
}

```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
all.pathways <- cellchat@netP$pathways
sprintf("Total number of pathways: %s", length(all.pathways))
print(all.pathways)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
interaction.on.pathways <- hash()
for (pathway in all.pathways){
  interaction.on.pathways[[pathway]] <- cellchat@netP$prob[, , pathway]
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
pathway.groups <- split(all.pathways, seq(1, 7))
all.clusters <- row.names(mat.weight)
```

# Visualization on pathway levels, sent signals 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
# for (cluster.id in all.clusters){
#   par(mfrow = c(3, 3))
#   cat(sprintf("### %s \n", cluster.id))
#   
#   for (pathway in all.pathways){
#     netVisual_aggregate(cellchat, signaling = c(pathway), layout = "circle", sources.use = c(cluster.id), targets.use = setdiff(row.names(mat.weight), c(cluster.id)), arrow.size = 1) 
#   }
#   cat("\n \n")
# }

for (cluster.id in all.clusters){
  cat(sprintf("## %s \n", cluster.id))
  cat("\n \n")
  for (group in names(pathway.groups)){
  cat(sprintf("### Group %s {.tabset}", group))
    cat("\n \n")
    for (pathway in pathway.groups[[group]]){
      cat(sprintf("#### Pathway: %s \n", pathway))
      pathways.show <- c(pathway) 
      par(mfrow = c(1,1), xpd=TRUE)
      netVisual_aggregate(cellchat, signaling = c(pathway), layout = "circle", sources.use = c(cluster.id), targets.use = setdiff(row.names(mat.weight), c(cluster.id)), arrow.size = 1) 
      cat("\n \n")
    }
}
}
```

# Visualization on pathway levels, received signals

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}

# for (cluster.id in all.clusters){
#   par(mfrow = c(3, 3))
#   cat(sprintf("### %s \n", cluster.id))
#   
#   for (pathway in all.pathways){
    # netVisual_aggregate(cellchat, signaling = c(pathway), layout = "circle", targets.use= c(cluster.id), sources.use = setdiff(row.names(mat.weight), c(cluster.id)), arrow.size = 1) 
#   }
#   cat("\n \n")
# }

for (cluster.id in all.clusters){
  cat(sprintf("## %s \n", cluster.id))
  for (group in names(pathway.groups)){
  cat(sprintf("### Group %s {.tabset} ", group))
    cat("\n \n")
    for (pathway in pathway.groups[[group]]){
      cat(sprintf("#### Pathway: %s \n", pathway))
      cat("\n \n")
      pathways.show <- c(pathway) 
      par(mfrow = c(1,1), xpd=TRUE)
      netVisual_aggregate(cellchat, signaling = c(pathway), layout = "circle", targets.use= c(cluster.id), sources.use = setdiff(row.names(mat.weight), c(cluster.id)), arrow.size = 1) 
      cat("\n \n")
    }
}
}

```

# Heatmap

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (group in names(pathway.groups)){
  cat(sprintf("## Group %s {.tabset} \n", group))
  cat("\n \n")
  for (pathway in pathway.groups[[group]]){
    cat(sprintf("### Pathway: %s \n", pathway))
    cat("\n \n")
    tryCatch(
        {
          p <- netVisual_heatmap(cellchat, signaling = c(pathway), color.heatmap = "Reds")
        },
        error=function(e) {
          print("error")
          p <- ggplot() + ggtitle("No data to generate heatmap for this pathway")
        }
    )
    print(p)
    cat("\n \n")
  }
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
```
