---
title: "Preliminary analysis, `r params$PROJECT`, Dataset: `r params$dataset.name`, sample: `r params$sample.id`"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: no
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: true
    theme: lumen
params:
  PROJECT: NA
  outdir: NA
  dataset.name: NA
  sample.id: NA
  input.outdir: NA
  project.type: NA
---
  
  
```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
$(document).ready(function() {
  $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
  // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>

**Note on the .html report: Click on the image to zoom-in and one more click to go back to the original**

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.height=10, fig.width=18}
##### clean up #####
# gc()
# rm(list = ls())

path.to.pipeline.src <- "/media/hieunguyen/HNSD01/src/src_pipeline/scRNA_GEX_pipeline_SeuratV5"

source(file.path(path.to.pipeline.src, "processes_src", "import_libraries.R"))
source(file.path(path.to.pipeline.src, "processes_src", "helper_functions.R"))

#####----------------------------------------------------------------------#####
# CONFIGURATIONS AND PREPRATIONS
#####----------------------------------------------------------------------#####
outdir <- params$outdir
PROJECT <- params$PROJECT
dataset.name <- params$dataset.name
sample.id <- params$sample.id
input.ourdir <- params$input.outdir
project.type <- params$project.type

# outdir <- "/media/hieunguyen/HNSD_mini/data/outdir"
# PROJECT <- "BrainMet_SeuratV5"
# dataset.name <- "GSE140393"
# sample.id <- "GSM4160577"
# input.outdir <- "/media/hieunguyen/HNHD01/outdir"
# project.type <- "Brain"

if (project.type == "Brain"){
  path.to.main.input <- file.path(input.outdir, 
                                sprintf("%s_%s_%s", PROJECT, dataset.name, sample.id), 
                                sprintf("%s_%s", dataset.name, sample.id)) 
  s.obj.raw <- readRDS(file.path(path.to.main.input, "s1_output", sprintf("%s_%s_%s.output.s1.rds", PROJECT, dataset.name, sample.id)))
  s.obj <- readRDS(file.path(path.to.main.input, "s8a_output", sprintf("%s_%s_%s.output.s8a.rds", PROJECT, dataset.name, sample.id)))
} else if (project.type == "Lung"){
  path.to.main.input <- file.path(input.outdir, 
                                sprintf("%s_%s", PROJECT, dataset.name), 
                                sprintf("%s", sample.id)) 
  s.obj.raw <- readRDS(file.path(path.to.main.input, "s1_output", sprintf("%s_%s.output.s1.rds", PROJECT, sample.id)))
  s.obj <- readRDS(file.path(path.to.main.input, "s8a_output", sprintf("%s_%s.output.s8a.rds", PROJECT, sample.id)))
}



s.obj <- subset(s.obj, Doublet_classifications == "Singlet")

path.to.main.output <- file.path(outdir, PROJECT, dataset.name, sample.id)
path.to.01.output <- file.path(path.to.main.output, "01_output")
dir.create(path.to.01.output, showWarnings = FALSE, recursive = TRUE)

random.gene <- sample(row.names(s.obj), 10)
count <- 0
for (i in random.gene){
  if (i == toupper(i)){
    count <- count + 1
  }
}
if (count >= 5){
  spec <- "Human"
} else {
  spec <- "Mouse"
}
topN <- 20
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.height=10, fig.width=18}
Idents(s.obj) <- "seurat_clusters"
if (file.exists(file.path(path.to.01.output, "DE_cluster_marker_genes.rds")) == FALSE){
  DefaultAssay(s.obj) <- "SCT"
  s.obj <- PrepSCTFindMarkers(s.obj)
  cluster.markers <- FindAllMarkers(object = s.obj, assay = "SCT", test.use = "wilcox", slot = "data", min.pct = 0.5)
  cluster.markers <- subset(cluster.markers, cluster.markers$p_val_adj < 0.05 & cluster.markers$avg_log2FC > 0)
  saveRDS(cluster.markers, file.path(path.to.01.output, "DE_cluster_marker_genes.rds"))
} else {
  cluster.markers <- readRDS(file.path(path.to.01.output, "DE_cluster_marker_genes.rds"))
}

```

# Preprocessing and QC for GEX data
## Raw data Quality control  {.tabset}

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (plot.name in names(s.obj@misc$all.QC)){
 
    cat('###',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$all.QC[plot.name])
  cat(' \n \n')
}
```

### Number of cells in each sample before filtering
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
table(s.obj.raw$name) %>% as.data.frame() %>% create_dt()
```

## Ambient RNA background correction 

### DecontX clusters {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (plot.name in names(s.obj@misc$ambient.cluster.RNA.plot)){
  
  cat('####',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$ambient.cluster.RNA.plot[plot.name])
  cat(' \n \n')
}
```

### Contamination level in each sample

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
ggplot(s.obj@meta.data, aes(x=name, y=AmbientRNA)) + 
  geom_boxplot()

```


### Ambient RNA contamination level {.tabset}

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (plot.name in names(s.obj@misc$ambient.contamination.plot)){
  
  cat('####',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$ambient.contamination.plot[plot.name])
  cat(' \n \n')
}
```

## Quality control after filtering {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
all.QC <- list()
  
  # Number of cells obtained per experiment/sample
  all.QC$cell.counts.plot <- ggplot(s.obj@meta.data, 
                                    aes(x=name , fill=name)) + 
    geom_bar() +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 14)) +
    ggtitle("Number of cells in each dataset")
  
  # distribution of number of UMI in each sample
  all.QC$nCountRNA.distribution <- ggplot(s.obj@meta.data,
                                          aes(color=name, x=nCount_RNA, fill = name)) + 
    geom_density(alpha = 0.2) +
    scale_x_log10() +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 14)) +
    ylab("Cell density") +
    geom_vline(xintercept = 500, color = "red") +
    ggtitle("Distribution of read depths in each sample")
  
  # distribution of number of features (genes detected) in each sample
  all.QC$nFeature_RNA.distribution <- ggplot(s.obj@meta.data,
                                             aes(color=name, x=nFeature_RNA, fill = name, y = ..scaled..)) + 
    geom_density(alpha = 0.2) +
    scale_x_log10() +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 14)) +
    ylab("Cell density") +
    geom_vline(xintercept = 1000, color = "red") +
    xlim(1000, 10000) +
    ggtitle("Distribution of number of detected genes in each sample")
  
  
  # scatter plot showing the relation between cell read-depth and number of genes detected.
  ## with Mitochondrial percentage
  all.QC$nCount.vs.nFeature.MT <- ggplot(s.obj@meta.data, 
                                         aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
    geom_point() + 
    scale_colour_gradient(low = "gray90", high = "black") +
    stat_smooth(method=lm, formula = y ~ x) + # apply a linear regression to show the relation, if existed.
    scale_x_log10() + 
    scale_y_log10() + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 12)) +
    facet_wrap(~name) +
    ggtitle("Scatter plot: nCount_RNA vs. nFeature_RNA, cmap % Mitochondrial genes")
  
  ## with Ribosome percentage
  all.QC$nCount.vs.nFeature.Ribo <- ggplot(s.obj@meta.data, 
                                           aes(x=nCount_RNA, y=nFeature_RNA, color=percent.ribo)) + 
    geom_point() + 
    scale_colour_gradient(low = "gray90", high = "black") +
    stat_smooth(method=lm, formula = y ~ x) + # apply a linear regression to show the relation, if existed.
    scale_x_log10() + 
    scale_y_log10() + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 12)) +
    facet_wrap(~name) +
    ggtitle("Scatter plot: nCount_RNA vs. nFeature_RNA, cmap % Ribosome genes")
  
  # Complexity: 
  
  ## We can see the samples where we sequenced each cell less have a higher overall complexity, 
  # that is because we have not started saturating the sequencing for any given gene for these samples. 
  # Outlier cells in these samples might be cells that have a less complex RNA species than other cells. 
  # Sometimes we can detect contamination with low complexity cell types like red blood cells via this metric. 
  # Generally, we expect the novelty score to be above 0.80. 
  
  ## More expanations: they are looking for cells that have a low number of genes with a high number of UMI counts. 
  # This likely means that you only captured transcripts from a low number of genes, and simply sequenced transcripts 
  # from those lower number of genes over and over again. This could be because of the cell type 
  # (such as a red blood cell having little to no RNA as they mentioned), or some other strange artifact.
  
  # Compute the complexity and add it to the s.obj@meta.data
  s.obj@meta.data <- s.obj@meta.data %>% 
    mutate(log10GenesPerUMI = log10(nFeature_RNA) / log10(nCount_RNA))
  
  all.QC$complexity <- ggplot(s.obj@meta.data,
                              aes(x=log10GenesPerUMI, color = name, fill=name)) +
    geom_density(alpha = 0.2) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    theme(plot.title = element_text(hjust=0.5, face="bold", size = 14)) +
    geom_vline(xintercept = 0.8) +
    ggtitle("Complexity: Log10(nCount_RNA) / log10(nFeature_RNA)")
  
  # add new slot for all.QC into the existed SEURAT OBJECT. 
  s.obj@misc$all.filtered.QC <- all.QC
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (plot.name in names(s.obj@misc$all.filtered.QC)){
 
    cat('###',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$all.filtered.QC[plot.name])
  cat(' \n \n')
}
```

### Number of cells in each samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
table(s.obj$name) %>% as.data.frame() %>% create_dt()
```


## Descriptive statistics and filtering threshold {.tabset}
This section is devoted to the descriptive statistics of the following varialbes: `nFeature_RNA, nCount_RNA, percent.mt, percent.ribo`. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (plot.item in c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo")){
  boxplot <- s.obj@meta.data %>% ggplot(aes_string(x = "name", y = plot.item)) +
    geom_boxplot(outlier.colour="black", outlier.shape=16,
                 outlier.size=2, notch=FALSE) +
    ggtitle(sprintf("Boxplot: Distribution of %s in each dataset", plot.item))
  cat('###', plot.item,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  
  print(boxplot)
  
  cat(' \n \n')
}
```



## Descriptive statistics + UMAP {.tabset}

### % Mitochondrial
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "percent.mt", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % Ribosome
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "percent.ribo", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % nCount RNA
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "nCount_RNA", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

### % nGenes 
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "nFeature_RNA", label = TRUE, label.size = 8, pt.size = 0.5, label.color = "red", )
```

# Dimension reduction with UMAP

## UMAP: all clusters 
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, label.size = 10, pt.size = 1, repel = TRUE) + 
  ggtitle(sprintf("UMAP: All clusters")) +
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```

## UMAP: all samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, label.size = 10, pt.size = 1, repel = TRUE, group.by = "name") + 
  ggtitle(sprintf("UMAP Sample: All samples after integrating")) +
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```

# Cluster marker genes
Identify differentially expressed genes in each cluster. 

## Feature plot {.tabset}
```{r echo=FALSE, fig.height=12, fig.width=20, message=FALSE, warning=FALSE, results='asis'}
for (cluster.id in sort(unique(cluster.markers$cluster))){
  DefaultAssay(s.obj) <- "SCT"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- FeaturePlot(object = s.obj, reduction = "RNA_UMAP", features = head(tmp.cluster.markers, 9)$gene, ncol = 3, label = TRUE, pt.size = 0.5, label.size = 5, label.color = "red", slot = "data")  
  print(p)
  cat("\n \n")
}
```

## Dot plot{.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (cluster.id in sort(unique(cluster.markers$cluster))){
  DefaultAssay(s.obj) <- "SCT"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- DotPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

## Violin plot{.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (cluster.id in sort(unique(cluster.markers$cluster))){
  DefaultAssay(s.obj) <- "SCT"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- VlnPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene, slot = "data")  
  print(p)
  cat("\n \n")
}
```

## Full tables of DE genes {.tabset}
```{r echo=FALSE, results='asis', include=FALSE}
cluster.markers %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
all.cluster.ids <- sort(unique(s.obj$seurat_clusters))
for (cluster.id in all.cluster.ids){
  tmp.table <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  cat(paste("\n\n### Cluster: ", cluster.id, "##\n"))
  print( htmltools::tagList(datatable(tmp.table, extensions = 'Buttons',
                                      filter = "top",
                                      options = list(dom = 'Blfrtip',
                                                     buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                                     lengthMenu = list(c(10,25,50,-1),
                                                                       c(10,25,50,"All")),
                                                     columnDefs = list(list(
                                                       targets = "_all",
                                                       render = JS(
                                                         "function(data, type, row, meta) {",
                                                         "return type === 'display' && data != null && data.length > 100 ?",
                                                         "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                                         "}")
                                                     ))
                                      ))))
  cat("\n \n")  
}
```

# Number of cells in each cluster in each sample
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
cell.count.in.clusters <- table(s.obj$name, s.obj$seurat_clusters) %>% as.data.frame() %>%
  pivot_wider(names_from = "Var2", values_from = "Freq") 
cell.count.in.clusters %>% create_dt()
```

# Cell type annotation by pathway analysis with CellMarkers 2.0
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (packageVersion("clusterProfiler") != "4.10.1"){
  remove.packages("clusterProfiler")
  remove.packages("DOSE")
  remove.packages("GOSemSim")
  path.to.install.dir <- "/media/hieunguyen/HD0/storage/offline_pkgs/clusterProfiler"
  install.packages(file.path(path.to.install.dir, "HDO.db_0.99.1.tar.gz"), type = "source", repos = NULL)
  install.packages(file.path(path.to.install.dir, "yulab.utils_0.1.4.tar.gz"), type = "source", repos = NULL)
  install.packages(file.path(path.to.install.dir, "GOSemSim_2.28.1.tar.gz"), type = "source", repos = NULL)
  install.packages(file.path(path.to.install.dir, "DOSE_3.28.2.tar.gz"), type = "source", repos = NULL) 
  install.packages(file.path(path.to.install.dir, "gson_0.1.0.tar.gz"), type = "source", repos = NULL)
  install.packages(file.path(path.to.install.dir, "clusterProfiler_4.10.1.tar.gz"), type = "source", repos = NULL) 
}

library(clusterProfiler)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if ("org.Mm.eg.db" %in% installed.packages() == FALSE){
  BiocManager::install("org.Mm.eg.db")
}
library("org.Mm.eg.db")

annotate_with_pathway <- function(cluster.id, topN = 20, spec = "Human"){
  if (topN == "all"){
      tmpdf <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  } else {
      tmpdf <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC)) %>% head(topN)    
  }
  if (nrow(tmpdf) != 0){
    path.to.cellmarker2.0 <- "/media/hieunguyen/HNHD01/data/CellMarkers_2.0"
    if (spec == "Human"){
      cell_marker_data <- readxl::read_excel(file.path(path.to.cellmarker2.0, "Cell_marker_Human.xlsx"))
      convertdf <- bitr(tmpdf$gene, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")    
    } else if (spec == "Mouse"){
      cell_marker_data <- readxl::read_excel(file.path(path.to.cellmarker2.0, "Cell_marker_Mouse.xlsx"))
      convertdf <- bitr(tmpdf$gene, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Mm.eg.db")   
    }
  
    cells <- cell_marker_data %>%
        dplyr::select(cell_name, GeneID) %>%
        dplyr::mutate(GeneID = strsplit(GeneID, ', ')) %>%
        tidyr::unnest()
    if (nrow(convertdf) > 2){
      x <- enricher(convertdf$ENTREZID, TERM2GENE = cells) %>% 
      as.data.frame() %>% subset(p.adjust <= 0.05)
      x <- x %>% subset(select = -c(geneID)) %>% arrange(desc(Count))      
    } else {
      x <- data.frame()
    }

  } else {
    x <- data.frame()
  }
  return(x)  
}

pathway.annotation <- hash()
```

## Result tables {.tabset}
```{r echo=FALSE, results='asis', include=FALSE}
annotate_with_pathway( cluster.id = 1, spec = spec, topN = topN) %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
annotationdf <- data.frame()
top5.annotationdf <- data.frame()

available.diff.clusters <- unique(cluster.markers$cluster)

for (cluster.id in all.cluster.ids){
  if (cluster.id %in% available.diff.clusters == FALSE){
    tmp.table <- data.frame(status = c("Cell type not detected!"))
  } else {
      tmp.table <- annotate_with_pathway(cluster.id = cluster.id, topN = topN, spec = spec)
  }

  if (nrow(tmp.table) == 0 | "status" %in% colnames(tmp.table)){
    tmp.table <- data.frame(status = c("Cell type not detected!"))
  } else {
      tmpdf <- data.frame(cluster.id = c(cluster.id), annotation = c(head(tmp.table, 1)$ID))
      annotationdf <- rbind(annotationdf, tmpdf)
      
      tmpdf <- data.frame(cluster.id = c(cluster.id), annotation = c(paste(head(tmp.table, 5)$ID, collapse = ", ")))
      top5.annotationdf <- rbind(top5.annotationdf, tmpdf)
    
  }
  
  cat(paste("\n\n### Cluster: ", cluster.id, "##\n"))
  print( htmltools::tagList(datatable(tmp.table, extensions = 'Buttons',
                                      filter = "top",
                                      options = list(dom = 'Blfrtip',
                                                     buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                                                     lengthMenu = list(c(10,25,50,-1),
                                                                       c(10,25,50,"All")),
                                                     columnDefs = list(list(
                                                       targets = "_all",
                                                       render = JS(
                                                         "function(data, type, row, meta) {",
                                                         "return type === 'display' && data != null && data.length > 100 ?",
                                                         "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                                         "}")
                                                     ))
                                      ))))
  cat("\n \n")  
}
```

## Top-1 annotation for each cluster
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
annotationdf %>% create_dt()
```

## Top-5 annotation for each clusters
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
top5.annotationdf %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
meta.data <- s.obj@meta.data %>% rownames_to_column("barcode") %>%
  rowwise() %>%
  mutate(prediction2 = ifelse(nrow(subset(annotationdf, annotationdf$cluster.id == seurat_clusters)) != 0, 
                              subset(annotationdf, annotationdf$cluster.id == seurat_clusters)$annotation,
                              "None")) %>%
  column_to_rownames("barcode")
meta.data <- meta.data[row.names(s.obj@meta.data),]
s.obj <- AddMetaData(s.obj, metadata = meta.data$prediction2, col.name = "prediction2")

DimPlot(object = s.obj, reduction = "RNA_UMAP", label.box = TRUE, label = TRUE, repel = TRUE, group.by = "prediction2")
```

## Number of cells in each annotated cluster (top-1 annotation)
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
table(s.obj$prediction2, s.obj$name) %>% as.data.frame() %>% pivot_wider(names_from = "Var1", values_from = "Freq") %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
if (file.exists(file.path(path.to.01.output,sprintf("%s_%s.rds", PROJECT, dataset.name, sample.id))) == FALSE){
  saveRDS(s.obj, file.path(path.to.01.output,sprintf("%s_%s.rds", PROJECT, dataset.name, sample.id)))  
} else {
  s.obj <- readRDS(file.path(path.to.01.output, sprintf("%s_%s.rds", PROJECT, dataset.name, sample.id)))
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
infodf <- data.frame(PROJECT = c(PROJECT), 
                     Dataset = c(dataset.name),
                     SampleID = c(sample.id),
                     Species = c(spec),
                     Num_cells = c(nrow(s.obj@meta.data)),
                     filename = c(sprintf("01_preliminary_analysis_%s_%s_%s.html", PROJECT, dataset.name, sample.id)))
writexl::write_xlsx(infodf, file.path(path.to.01.output, "infodf.xlsx"))
```


